---
title: "cusecomp"
output: html_document
---


```{r}
library(nflfastR)
library(tidyverse)
scheme_data <- read_csv("scheme_data.csv")
pbp_data <- nflfastR::load_pbp(2021)

# Edit team names for later merging
scheme_data <- scheme_data %>%
  mutate(posteam = case_when(
    offense %in% c("ATL", "BUF", "CAR", "CHI", "CIN", "DAL", "DEN", "DET", "GB", "IND", "JAX", "KC", "LA", "LAC", "LV", "MIA", "MIN", "NE", "NO", "NYG", "NYJ", "PHI", "PIT", "SEA", "SF", "TB", "TEN", "WAS") ~ offense,
    offense == "ARZ" ~ "ARI",
    offense == "BLT"~ "BAL",
    offense == "CLV" ~ "CLE",
    offense == "HST" ~ "HOU",
    offense == "OAK" ~ "LV",
    offense == "SD" ~ "LAC",
    offense == "SL" ~ "LA"
  ),
  defteam = case_when(
    defense %in% c("ATL", "BUF", "CAR", "CHI", "CIN", "DAL", "DEN", "DET", "GB", "IND", "JAX", "KC", "LA", "LAC", "LV", "MIA", "MIN", "NE", "NO", "NYG", "NYJ", "PHI", "PIT", "SEA", "SF", "TB", "TEN", "WAS") ~ defense,
    defense == "ARZ" ~ "ARI",
    defense == "BLT"~ "BAL",
    defense == "CLV" ~ "CLE",
    defense == "HST" ~ "HOU",
    defense == "OAK" ~ "LV",
    defense == "SD" ~ "LAC",
    defense == "SL" ~ "LA"
  )) %>%
  rename(yardline_100 = yards_to_go, ydstogo = distance, qtr = quarter, quarter_seconds_remaining = seconds_left_in_quarter) %>%
  mutate(game_seconds_remaining = quarter_seconds_remaining+(4-qtr)*900,
         week = case_when(
           week < 19 ~ week,
           week == 28 ~ 19,
           week == 29 ~ 20,
           week == 30 ~ 21,
           week == 32 ~ 22
         )) %>%
  filter(down > 0, rps %in% c("R", "P")) %>%
  group_by(season, week, posteam, defteam, qtr, down, ydstogo, yardline_100) %>%
  arrange(desc(game_seconds_remaining)) %>%
  mutate(dupe_number = 1:n()) %>%
  ungroup()

pbp_data <- pbp_data %>%
  filter(down > 0) %>%
  ungroup() %>%
  select(season, week, posteam, defteam, yardline_100, quarter_seconds_remaining:game_seconds_remaining, qtr:goal_to_go, ydstogo, yards_gained, shotgun:run_gap, ep, epa, wp, wpa, qb_hit:sack, passer, rusher, receiver, pass, rush) %>%
  group_by(season, week, posteam, defteam, qtr, down, ydstogo, yardline_100) %>%
  arrange(desc(game_seconds_remaining)) %>%
  mutate(dupe_number = 1:n()) %>%
  ungroup()

# Use posteam, defteam, and other play information to merge scheme_data with pbp_data
merged_data <- left_join(scheme_data, pbp_data, by = c("season", "week", "posteam", "defteam", "qtr", "down", "ydstogo", "yardline_100", "dupe_number")) %>%
  filter(down > 0, rps %in% c("R", "P"))



kcepa <- merged_data %>%
  group_by(offense) %>%
  subset(offense == "KC")


topoffense <- kcepa %>%
  filter(offense_personnel%in%c("3-0-1-1", "2-1-1-1", "2-0-1-2", "1-0-1-3"))



merged_data <- merged_data %>%
  mutate(score_diff = off_score_before - def_score_before)

kcepa %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n(), run_rate = mean(rush)) %>%
  arrange(desc(mean_epa)) %>%
  ungroup()

kcepa%>% #important traits of KC offense by personnel
  filter(rush == 0) %>%
  #filter(!(pass_direction %in% c("X", NA))) %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n(), target_depth = mean(target_depth), play_action = mean(play_action), time_throw = mean(time_to_throw), time_pressure = mean(time_to_pressure), yac = mean(yards_after_catch, na.rm = TRUE), mean_yards_to_go = mean(yardline_100), mean_down = mean(down)) %>%
  arrange(desc(mean_epa)) %>%
  ungroup() 


topoffense %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = down, fill = rush)) +
  geom_bar(position = "fill", color = "black") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  facet_wrap(~offense_personnel) +
  labs(x = "Down")+
   scale_fill_manual(values=c("#E31837",
                             "#FFB81C"))


topoffense %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = down, fill = offense_personnel)) +
  geom_bar(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Down", y = "Count") +
  facet_wrap(~offense_personnel)+
scale_fill_manual(values=c("#E31837",
                             "#FFB81C", "black", "grey"))



```




```{r}
topoffense %>%
  ggplot(aes(x= offense_personnel, fill = rps)) +
  geom_bar(position = "fill", color = "black")+
  labs(x = "Offensive Personnel", y="Frequency")+
  ggtitle("Frequency of Pass/Rush in the Chiefs Top Offensive Formations")+
   scale_fill_manual(values=c("#E31837",
                             "#FFB81C"))
merged_data %>% #epa/play for passing plays by defensive personnel
  filter(posteam == "KC") %>%
  group_by(defense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n()) %>%
  arrange(desc(mean_epa)) %>%
  filter(count >= 20) %>%
  ungroup()


merged_data %>% #EPA by offensive personnel type, passing 
  filter(posteam == "KC" & rush == 0) %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n()) %>%
  arrange(desc(mean_epa)) %>%
  ungroup() %>%
  filter(count >=30)

merged_data %>% #epa/play for passing plays by defensive personnel
  filter(posteam == "KC" & rush == 0) %>%
  group_by(defense_personnel) %>%
  summarize(mean_passepa = mean(epa), count = n()) %>%
  arrange(desc(mean_passepa)) %>%
  filter(count >= 20) %>%
  ungroup()

merged_data %>% #EPA by offensive personnel type, passing 
  filter(posteam == "KC" & rush == 0) %>%
  group_by(offense_personnel) %>%
  summarize(mean_passepa = mean(epa), count = n()) %>%
  arrange(desc(mean_passepa)) %>%
  ungroup() %>%
  filter(count >=30)



merged_data %>% #KC epa/play by number of box players for passing plays
  filter(posteam == "KC" & rush == 0) %>%
  group_by(box_players) %>%
  summarize(mean_passepa = mean(epa), count = n(), target_depth = mean(target_depth), time_to_pass = mean(time_to_throw)) %>%
  arrange(desc(mean_passepa)) %>%
  ungroup()

merged_data %>%
  filter(posteam == "KC") %>%
  summarize(pass_ratio = mean(rush)) #Chiefs run about 30% of the time


merged_data %>%
  filter(posteam == "KC") %>%
  group_by(defense_personnel, offense_personnel) %>%
  summarize(mean_epa = mean(epa), pass_ratio = 1 - mean(rush), count = n()) %>%
  arrange(desc(mean_epa)) %>%
  filter(count >= 20) %>%
  ungroup()

merged_data %>%
  filter(posteam == "KC" & rush == 0) %>%
  group_by(defense_personnel, offense_personnel) %>%
  summarize(mean_epa = mean(epa), pass_ratio = 1 - mean(rush), count = n(), play_action = mean(play_action)) %>%
  arrange(desc(mean_epa)) %>%
  filter(count >= 20) %>%
  ungroup()

merged_data %>%
  filter(posteam == "KC") %>%
  group_by(play_action) %>%
  summarize(mean_epa = mean(epa))


kcoffense %>%
  ggplot(aes(x = target_depth, y = epa, color = offense_personnel)) +
  geom_point()


table1 <- merged_data %>%
  filter(posteam == "KC") %>%
  group_by(offense_personnel, defense_personnel) %>%
  summarize(mean_epa = mean(epa), pass_ratio = 1 - mean(rush), count = n(), play_action = mean(play_action), blitz_ratio = mean(blitz), box = mean(box_players), down_m = mean(down), time_to_pres = mean(time_to_pressure)) %>%
  arrange(desc(play_action)) %>%
  filter(count >= 20) %>%
  ungroup()
write_csv(table1, file = "table1.csv")
```



```{r}

topoffense %>%
  filter(pass_direction != "NA")%>%
  filter(pass_direction != "X")%>%
  ggplot(aes(x= offense_personnel, fill = pass_direction)) +
  geom_bar(position = "fill", color="Black")+
  labs(x = "Offensive Personnel", y="Frequency")+
  ggtitle("Frequency of Pass Direction in the Chiefs Top Offensive Formations ")+
   scale_fill_manual(values=c("#E31837","white",
                             "#FFB81C"))
target_offense <- c("2-0-1-2", "2-1-1-1", "3-0-1-1")

merged_data %>%
  filter(posteam == "KC") %>%
  filter(offense_personnel == target_offense) %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = offense_personnel, fill = rush)) +
  geom_bar(position = "dodge", color = "Black") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Offensive Personnel", y = "Count")+
   scale_fill_manual(values=c("#E31837","#FFB81C"))+
  ggtitle("Number of Passing/Rushing plays per Chiefs Offensive Personnel")


target_defense = c("1/4/06", "2/3/06", "2/4/05", "3/2/06", "3/3/05", "3/4/04", "4/1/06", "4/2/05", "4/3/04")
merged_data %>% ## need to fix
  filter(posteam == "KC") %>%
  filter(defense_personnel == target_defense) %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = defense_personnel, fill = rush)) +
  geom_bar(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Defensive Personnel", y = "Count")+
   scale_fill_manual(values=c("#E31837","#FFB81C"))+
  ggtitle("Number of Passing/Rushing plays per Chiefs Opponent's Defensive Personnel")


merged_data %>% ## need to fix
  filter(posteam == "KC") %>%
  filter(offense_personnel == target_offense) %>%
  mutate(play_a = factor(play_action)) %>%
  ggplot(aes(x = offense_personnel, fill = play_a)) +
  geom_bar(position = position_fill(), color = "black") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Offensive Personnel", y = "Proportion")+
   scale_fill_manual(values=c("#E31837","#FFB81C"))+
  ggtitle("Proportion of Play Action Plays in Chiefs Top Offensive Personnel")


merged_data %>%
  filter(posteam == "KC") %>%
  filter(offense_personnel%in%c("2-0-1-2", "2-1-1-1", "3-0-1-1"))%>%
  ggplot(aes(x = epa)) +
  geom_histogram() +
  geom_vline(xintercept = 0, linetype="dotted", 
                color = "blue", size=1.5)+
  facet_grid(~offense_personnel) +
  labs(x = "epa/play", y = "offensive scheme")+
  theme(axis.text.x = element_text(angle = 60, hjust=1)) 



merged_data %>%
  filter(posteam == "KC") %>%
  group_by(blitz, offense_personnel) %>%
  summarize(mean_epa = mean(epa), count=n())%>%
  filter(count > 30)



merged_data %>%
  filter(posteam == "KC") %>%
  group_by(blitz, defense_personnel) %>%
  summarize(mean_epa = mean(epa), count=n())%>%
  filter(count > 30)

merged_data %>%
  filter(posteam == "KC") %>%
  group_by(blitz) %>%
  summarize(mean_epa = mean(epa), count=n())%>%
  filter(count > 30)


```


```{r}

topoffense %>%
  ggplot(aes(x= offense_personnel, fill = mofo_coverage_played)) +
  geom_bar(position = "fill", color = "black")+  
  labs(x = "Offensive Personnel", y="Frequency")+
   ggtitle("Frequency of Middle of Field Coverage in \nthe Chiefs Top Offensive Formations ")+
   scale_fill_manual(values=c("#E31837",
                             "#FFB81C"))


topoffense %>%
  ggplot(aes(x= offense_personnel, fill = mofo_coverage_shown)) +
  geom_bar(position = "fill", color = "black")+
  labs(x = "Offensive Personnel", y="Frequency")+
   ggtitle("Frequency of Middle of Field Coverage Shown in \nthe Chiefs Top Offensive Formations ")+
   scale_fill_manual(values=c("#E31837",
                             "#FFB81C"))


topoffense%>%
  group_by(mofo_coverage_shown, mofo_coverage_played)%>%
  summarize(meanepa = mean(epa))




\
```
```{r}
otheroffense <- merged_data %>%
  group_by(offense) %>%
  subset(offense != "KC")%>%
  filter(offense_personnel%in%c("3-0-1-1", "2-1-1-1", "2-0-1-2", "1-0-1-3"))

otheroffense %>%
group_by(offense_personnel, down)%>%
summarize(mean_epa = mean(epa), count=n())

topoffense %>%
  group_by(offense_personnel)%>%
summarize(mean_epa = mean(epa), count=n())
  
```
```{r}
  kcepa %>%
  ungroup() %>%
    group_by(defense_personnel)%>%
    summarize(mean_epa = mean(epa), count = n())%>%
    filter(count >30)


 kcepa %>%
  ungroup() %>%
    group_by(coverage_scheme)%>%
    summarize(mean_epa = mean(epa), count = n())%>%
    filter(count >30)
  
  
 kcepa %>%
    group_by(coverage_scheme)%>%
    count()
  
  kcepa %>%
    group_by(num_pass_rush_players)%>%
    summarize(mean = mean(epa), count = n())
  
  kcepa %>%
    group_by(box_players)%>%
    summarize(mean = mean(epa), count = n())
  
  kcepa%>%
  ggplot (aes(x=num_pass_rush_players, y=epa)) +
    geom_violin()
  
```
```{r}
defense <- kcepa %>%
  filter(defense_personnel%in%c("2/4/05", "2/3/06", "3/3/05", "4/1/06", "4/2/05"))

defense%>%
  group_by(defense_personnel)%>%
  summarize(mean = mean(epa), count=n())


defense%>%
  ggplot(aes(x=defense_personnel, fill=rps))+
  geom_bar(position="dodge")

defense%>%
   filter(offense_personnel%in%c("3-0-1-1", "2-1-1-1", "2-0-1-2", "1-0-1-3"))%>%
  ggplot(aes(x=defense_personnel, fill=offense_personnel))+
  geom_bar(position="fill")

defense%>%
  mutate(blitz=factor(blitz))%>%
  ggplot(aes(x=defense_personnel, fill=blitz))+
  geom_bar(position="fill")

defense%>%
  mutate(down=factor(down))%>%
  ggplot(aes(x=defense_personnel, fill=down))+
  geom_bar(position="dodge")



defense%>%
  group_by(defense_personnel)%>%
  filter(defense_personnel%in%c("2/4/05", "2/3/06", "3/3/05", "4/1/06", "4/2/05"))%>%
  summarize(mean_epa = mean(epa), count=n(), mean_box = mean(box_players), mean_yards = mean(ydstogo), mean_time = mean(time_to_throw))


```

```{r}
opponents <- merged_data%>%
  filter(defense%in%c("MIA", "NE", "NYJ", "CHI", "IND", "HOU", "JAX", "DET", "ATL", "CAR", "NO", "TB", "LA", "SEA", "MIN", "SF", "ARI"))

opponents %>%
   filter(offense_personnel%in%c("3-0-1-1"))%>%
  group_by(defense)%>%
  summarize(meanepa = mean(epa), count=n())%>%
  
```
```{r}
neworleans <- opponents%>%
  filter(defense%in%c("NO"))


neworleans%>%
  group_by(offense_personnel)%>%
  summarize(meanepa = mean(epa), count=n())%>%
  filter(count>30)


neworleans%>%
   mutate(down=factor(down))%>%
  filter(defense_personnel%in%c("3/2/06","4/2/05"))%>%
  ggplot(aes(x=defense_personnel, fill = down))+
  geom_bar()




larams <- opponents%>%
  filter(defense%in%c("LA"))

larams%>%
  group_by(offense_personnel)%>%
  summarize(meanepa = mean(epa), count=n())%>%
  filter(count > 30)


larams%>%
   mutate(down=factor(down))%>%
  filter(defense_personnel%in%c("2/3/06","2/4/05", "3/3/05", "3/4/04"))%>%
  filter(offense_personnel%in%c("3-0-1-1"))%>%
  group_by(defense_personnel)%>%
  summarize(meanepa = mean(epa), count=n())
```
```{r}
merged_data %>%
  filter(posteam == "KC") %>%
  filter(offense_personnel == target_offense) %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = offense_personnel, fill = rush)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Offensive Personnel", y = "Proportion")

merged_data %>% ## need to fix
  filter(posteam == "KC") %>%
  filter(offense_personnel == target_offense) %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = offense_personnel, fill = rush)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Offensive Personnel", y = "Proportion")

merged_data %>% ## need to fix
  filter(posteam == "KC") %>%
  filter(defense_personnel == target_defense) %>%
  mutate(rush = factor(rush)) %>%
  ggplot(aes(x = offense_personnel, fill = rush)) +
  geom_bar(position = position_fill()) +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Defensive Personnel", y = "Proportion")

merged_data %>% ## need to fix
  filter(posteam == "KC") %>%
  filter(offense_personnel == target_offense) %>%
  mutate(play_a = factor(play_action)) %>%
  ggplot(aes(x = offense_personnel, fill = play_a)) +
  geom_bar(position = position_fill()) +
  theme(axis.text.x = element_text(angle = 60, hjust=1)) +
  labs(x = "Defensive Personnel", y = "Proportion")

merged_data %>%
  filter(posteam == "KC") %>%
  ggplot(aes(x = epa)) +
  geom_histogram() +
  facet_grid(~offense_personnel) +
  labs(x = "epa/play", y = "offensive scheme")
  theme(axis.text.x = element_text(angle = 60, hjust=1)) 
  
merged_data %>%
  filter(posteam == "KC") %>%
  ggplot(aes(x = epa)) +
  geom_histogram() +
  facet_wrap(~defense_personnel) +
  labs(x = "epa/play", y = "offensive scheme")
  theme(axis.text.x = element_text(angle = 60, hjust=1))
```
```{r}
kcepa <- kcepa %>%
  mutate(score_diff = off_score_before - def_score_before)

kcepa %>%
  filter(yardline_100 <= 90 & yardline_100 >= 60) %>%
  filter(offense_personnel == "3-0-1-1") %>%
  filter(down == 1 & abs(score_diff) < 8) %>%
  filter(screen == 0) %>%
  summarize(down_yards = mean(ydstogo), mean_rush = mean(rush), mean_epa = mean(epa), count = n(), target_depth = mean(target_depth), play_action = mean(play_action), time_throw = mean(time_to_throw), time_pressure = mean(time_to_pressure), yac = mean(yards_after_catch, na.rm = TRUE), mean_yards_to_go = mean(yardline_100), mean_down = mean(down), m_screen = mean(screen), ypp = mean(yards_gained)) %>%
  arrange(desc(mean_epa)) %>%
  ungroup()


kcepa %>% #important traits of KC offense by personnel
  filter(rush == 0) %>%
  #filter(!(pass_direction %in% c("X", NA))) %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n(), target_depth = mean(target_depth), play_action = mean(play_action), time_throw = mean(time_to_throw), time_pressure = mean(time_to_pressure), yac = mean(yards_after_catch, na.rm = TRUE), mean_yards_to_go = mean(yardline_100), mean_down = mean(down)) %>%
  arrange(desc(mean_epa)) %>%
  filter(count > 25)%>%
  ungroup()

```


```{r}
kcoffense %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n(), run_rate = mean(rush)) %>%
  arrange(desc(mean_epa)) %>%
  ungroup()

kcoffense %>% #important traits of KC offense by personnel
  filter(rush == 0) %>%
  #filter(!(pass_direction %in% c("X", NA))) %>%
  group_by(offense_personnel) %>%
  summarize(mean_epa = mean(epa), count = n(), target_depth = mean(target_depth), play_action = mean(play_action), time_throw = mean(time_to_throw), time_pressure = mean(time_to_pressure), yac = mean(yards_after_catch, na.rm = TRUE), mean_yards_to_go = mean(yardline_100), mean_down = mean(down)) %>%
  arrange(desc(mean_epa)) %>%
  ungroup()


